Surface-driven registration method for the structure-informed segmentation of diffusion MR images\tnoteref{t1}

### 2015-10-26 Publications/2015-NeuroImage/02-abstract.tex
new abstract!

% -*- root: 00-main.tex -*-
\newcomment[RV\#1(C.1)]{%
Current methods using diffusion MRI to map the microstructure and connectivity of the
  human brain increasingly require precise delineations of anatomical structures.
The problem is typically solved by projecting the structural information extracted
  in T1-weighted images, by means of nonlinear registration since diffusion
  MRI data present geometrical distortions.
Here we propose \regseg{}, which segments homogeneous regions within multivariate
  images by projecting a set of nested surfaces extracted from structural images.}
The surfaces evolve through a free-form deformation field supported by the B-spline basis
  to optimally map the contours onto the data in the target space.
We tested the functionality of \regseg{} using four digital phantoms warped with known and
  randomly generated deformations, where subvoxel accuracy was achieved.
\newcomment[RV\#2(C.11)]{%
We proposed a set of surfaces that define a segmentation model of the fractional anisotropy and
  the apparent diffusion coefficient maps derived from diffusion MRI datasets.}
These surfaces extracted from the same-subject T1-weighted image successfully segmented 16 real
  diffusion MRI datasets from the Human Connectome Project that were warped by realistic and nonlinear
  distortions.
We computed the misregistration error of the surfaces estimated by \regseg{} with respect
  to their theoretical location using the ground truth, thereby obtaining a 95\% CI of 0.56--0.66 mm
  distance between corresponding mesh vertices, which was below the 1.25 mm isotropic resolution of the images.
We also compared the performance of the proposed method with a widely used registration tool, which showed
  that \regseg{} outperformed this method in our settings.

>>>>

% -*- root: 00-main.tex -*-
\revcomment[RV\#1(C.1)\par RV\#2(C.10)]{%
Current methods for processing \gls*{dmri} to map the connectivity of the human brain
  require precise delineations of anatomical structures.
Image segmentation on the native \gls*{dmri} with sub-pixel accuracy has been unsuccessful,
  and the projection of the structural information from \gls*{t1} images is challenging
  since \gls*{dmri} data present geometrical distortions.
To solve this problem, we propose \regseg{}, a surfaces-to-volume nonlinear registration method
  that segments homogeneous regions within multivariate images by projecting a set of nested
  reference-surfaces.}
We first verify the subpixel accuracy of \regseg{} on digital phantoms,
\revcomment[RV\#2(C.11)]{%
and then we evaluate the tool in the aforementioned application.
We propose a set of surfaces extracted from the \gls*{t1} image to successfully segment
  the bivariate volume comprehending the \gls*{fa} and the \gls*{adc} maps derived from the
  corresponding \gls*{dmri} dataset of the same subject.}
We present an evaluation framework, in which the target \gls*{dmri} are deformed with
  magnetic susceptibility -derived and realistic distortions.
Using the framework, we analyze the misregistration error of the surfaces estimated by \regseg{}
  on 16 real datasets of the Human Connectome Project, thereby obtaining an error distribution
  with 95\% CI of 0.56--0.66 mm, below the \gls*{dmri} resolution (1.25 mm, isotropic).
Finally, we cross-compare \regseg{} against a widely-used volume-to-volume registration method,
  and we obtain a significantly lower misregistration error with the proposed tool.

### 2015-10-26 Publications/2015-NeuroImage/04-methods.tex
new abstract!

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:regseg-bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
For convenience and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions \citep{esteban_mbis_2014} with parameters
  $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
\revcomment[RV\#1(C.4)]{%
Even though the features being segmented are not generally i.i.d., the spatial interdependency of
  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
  multivariate normal distributions.} Therefore,

>>>>

  \begin{equation}
  P(M \mid U) = \underset{l}{\prod} \underset{\vec{r}\in \Omega_l}{\prod}
    P\left( \vec{f}' \mid U \right),
  \label{eq:regseg-bayes_aposteriori}
  \end{equation}
  where $\vec{f}' = M(\vec{r}')$ is the feature vector at the displaced
  position $\vec{r}'$ \eqref{eq:regseg-transform} in the moving image.
For convenience and because it has been shown to be an appropriate approximation
  \citep{leemput_automated_1999a,cuadra_comparison_2005}, we introduce two assumptions for each
  region $\Omega_l$:
  1) the features are i.i.d.; and 2) they can be modeled by multivariate normal
  distributions \citep{esteban_mbis_2014} with parameters
  $\lbrace \boldsymbol{\mu}_l, \boldsymbol{\Sigma}_{l} \rbrace$
  for each region $\Omega_l$.
\revcomment[RV\#1(C.4)]{%
Even though the features being segmented are not generally i.i.d., the spatial interdependency of
  voxels is implicitly supported by the piecewise smooth partition of the space \omegaset{}.
In \autoref{fig:regseg-model} it is shown that marginal distributions of data can be approximated by
  multivariate normal distributions.} Therefore,

### 2015-10-26 Publications/2015-NeuroImage/07-conclusion.tex
new abstract!

\section*{Availability and reproducibility statement}
\label{sec:regseg-availability}
We considered the reproducibility of our results as a design requirement.
Therefore, we used real data obtained from a publicly available repository
  (the Human Connectome Project \citep{essen_human_2012}) and all of the software
  utilized in this study is also publicly available.
\Regseg{} was implemented on top of ITK-4.6 (Insight Registration and 
  Segmentation Toolkit, \url{http://www.itk.org}).
The evaluation instruments (\autoref{fig:regseg-evworkflows}) were implemented using
  \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
  are publicly available under a unique package \citep{esteban_acweregistration_2015}.

>>>>

\section*{Availability and reproducibility statement}
\label{sec:regseg-availability}
We considered the reproducibility of our results as a design requirement.
Therefore, we used real data obtained from the Human Connectome Project \citep{essen_human_2012}
  and all of the software utilized in this study is also publicly available.
\Regseg{} was implemented on top of ITK-4.6 (Insight Registration and
  Segmentation Toolkit, \url{http://www.itk.org}).
The evaluation instruments (\autoref{fig:regseg-evworkflows}) were implemented using
  \emph{nipype} \citep{gorgolewski_nipype_2011} to assess their reproducibility.
All of the research elements (data, source code, figures, manuscript sources, etc.) involved in this study
  are publicly available under a unique package \citep{esteban_acweregistration_2015}.
